Polynomial phase estimation by phase unwrapping

### 2012-10-22 paper.tex
edits to paper. Removed the symmetrisation proof, which was very close to Pollards

@@

% AOS,AOAS: If there are supplements please fill:
%\begin{supplement}[id=suppA]
%  \sname{Supplement A}
%  \stitle{Title}
%  \slink[url]{http://lib.stat.cmu.edu/aoas/???/???}
%  \sdescription{Some text}
%\end{supplement}

@@

An obvious estimator of the unknown coefficients is the least squares estimator.  
%This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  
When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the least squares estimator is an effective approach, being both computationally efficient and statistically accurate~\cite{Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}. When $m \geq 2$ the computational complexity of the least squares estimator is large, and for this reason authors have considered alternative estimators. These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995} and the high order phase function~\cite{Farquharson_another_poly_est_2005,Porat_asympt_HAF_DPT_1996}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.

@@

An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $N^{-k}$.  This is perhaps not surprising, since it is the same rate observed in polynomial regression.  However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.  For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$.  %This proof technique appears to be novel, and may be useful in other problems.  
The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.  Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.  We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.  Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.

@@

A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^d \}
\]
where $\Bbf \in \reals^{n \times d}$ is an $n \times d$ matrix of rank $d$, called the generator matrix.  If $n = d$ the lattice is said to be full rank.  Lattices are discrete Abelian groups under vector addition.  They are subgroups of the Euclidean group $\reals^n$.  Lattices naturally give rise to tessellations of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.  One choice for a set of coset representatives is a fundamental parallelepiped; the parallelepiped generated by the columns of a generator matrix.  Another choice is based on the Voronoi cell, i.e., those points from $\reals^n$ nearest (with respect to the Euclidean norm here) to the lattice point at the origin.  It is always possible to construct a rectangular set of representatives, as the next proposition will show.  We will use these rectangular regions for describing the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.  These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.

@@

In general there are infinitely many choices for the identifiable region. A natural choice is the Voronoi cell of $L_{m+1}$ used in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}. Another potential choice is a fundamental parallelepiped of $L_{m+1}$. In this paper we will use the rectangular set constructed using Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region}. Observe that $\Pbf$ is upper triangular with $k$th diagonal element equal to $\tfrac{1}{k!}$.  So this rectangular region is
\begin{equation}\label{eq:rectangular_identifiable_region}
B = \prod_{k=0}^{m}\left[ -\frac{0.5}{k!}, \frac{0.5}{k!}  \right).
\end{equation}
We will make use of this region when deriving the statistical properties of the LSU estimator in the next section.

@@

Let $W = \{1,2,\dots, N\}$ and let $K$ be a subset of $W$.  For any integer $h$, let
\begin{equation} \label{eq:S(h,G)def} 
A(h,K) = \big\{ n \mid n + ih \in K \;\forall\; i \in \{0,1,\dots,m\} \big\}
\end{equation}
be the set containing all integers $n$ such that the arithmetic progression
\[
n, \,\, n + h, \,\, n + 2h, \,\, \dots, \,\, n + mh
\]
of length $m+1$ is contained in the subset $K$.  If $K$ is a small subset of $W$ then $A(h,K)$ might be empty. However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty. Denote by $K \backslash \{r\}$ the set $K$ with the element $r$ removed.

@@

During the proof of strong consistency we made use of the fact that
\begin{equation}~\label{eq:SNVNunifmlln2}
\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0
\end{equation}
almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here.  In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$. Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.  Now $\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty$ for any $\epsilon > 0$ by Lemma~\ref{lem:vn}, and~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.

\begin{lemma} \label{lem:vn} $\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 2$.
 \end{lemma}
\begin{proof}
Consider a rectangular grid of points spaced over the identifiable region $B$.  We use $\lambdabf[\rbf]$, where $\rbf \in \ints^{m+1}$, to denote the grid point 
 %\begin{align*}
\[ 
%\lambdabf[\rbf] = &\left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \right.\\ 
%&\dots, \; \frac{r_k}{N^{b+k}} - \frac{1}{2(k!)}, \; \dots \\
%&\left. \dots, \; \frac{r_m}{N^{b+m}} - \frac{1}{2(m!) }\right]
 \lambdabf[\rbf] = \left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \; \dots, \; \frac{r_m}{m!N^{b+m}} - \frac{1}{2(m!) }\right]
\]
 %\end{align*}
 for some constant $b>0$.  Adjacent grid points are separated by $\tfrac{1}{N^b}$ in the zeroth coordinate, $\tfrac{1}{N^{b+1}}$ in the first coordinate and $\tfrac{1}{k!N^{b+k}}$ in the $k$th coordinate. Let
 \[
 B[\rbf]=\left\{  \xbf\in\reals^{m+1}  ; \frac{r_k}{N^{b+k}}\leq x_k + \frac{1}{2(k!)} < \frac{r_k + 1}{N^{b+k}} \right\}  .
 \]
and let $G$ be the finite set of grid points
\[
G = \left\{ \xbf \in \ints^{m+1} \mid x_k = 0,1,2\dots,N^{b+k}-1  \right\}.
\]
The total number of grid points is $|G| = N^{(m+1)(2b + m)/2}$, and the $B[\rbf]$ partition $B$, that is, $B = \cup_{\rbf \in G}B[\rbf]$.  Now 
\begin{align}
 \sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  &= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
 &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
 \end{align}
From Lemma~\ref{lem:supVjk} it will follow that
\[
\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})
\]
for any $\epsilon > 0$ and $c < 2$.  In Lemma~\ref{lem:supBVn} we show that
\[
\sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\abs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])} < 2\frac{m+1}{N^b}.
\]
%that is, $D_N$ does not change much within any of the regions $B[\rbf]$ neighbouring the grid points.  
Combining these results with~\eqref{eq:gridanddensespace}, we obtain
\[
\prob\left(\sup_{\lambdabf \in B }\sabs{ D_N(\lambdabf)} > \frac{\epsilon}{2} +  \frac{2(m+1)}{N^b} \right) = O(e^{-c \epsilon^2 N}),
\]
and for sufficiently large $N$, we have $\epsilon/2 + \frac{2(m+1)}{N^b} < \epsilon$ completing the proof.  It remains to prove Lemmas~\ref{lem:supVjk} and~\ref{lem:supBVn}.
\end{proof}

@@

\begin{lemma}\label{lem:supVjk}
$\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 8$.
\end{lemma}
 \begin{proof}
Fix $\lambdabf$ and write $D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n}$, where
 \[
 Z_{n}=\fracpart{  \Phi_n+\sum_{k = 0}^{m}{\lambda_k n^k} }^{2} - \expect \fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }^{2}
 \]
 are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,%From Lemma~\ref{lem:zero_mean_indepent_sum_bound},
 \[
  \prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N},
 \]
and so,
 \begin{align*}
 \prob\cubr{  \sup_{\rbf\in G}\left\vert D_N(\lambdabf[\rbf]) \right\vert >\epsilon }  &\leq \sum_{\rbf \in G} \prob\scubr{  \left\vert D_N(\lambdabf[\rbf])  \right\vert >\epsilon } \nonumber \\
&= 2 |G| e^{ - 8 \epsilon^2 N} = O(e^{-c \epsilon^2 N}), \label{eq:betaprobconv}
\end{align*}
where $c$ is any real number less than $8$, since $|G| = N^{(m+1)(2b + m)/2}$ is polynomial in $N$.
\end{proof}

@@

\begin{lemma}\label{lem:boundedsquarefracparts}
$\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|$ for all real numbers $x$ and $\delta$.
\end{lemma}
\begin{proof}
Since $\abs{\delta} \leq \abs{n + \delta}$ for all $\delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$ and $n \in \ints$, the result will follow if we can show that it holds when both $x$ and $\delta$ are in $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.  Also, for reasons of symmetry, we need only show that it holds when $\delta \geq 0$.  Now
\[
\fracpart{x + \delta}^2 - x^2 = \begin{cases}
2x\delta + \delta^2, & x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta) \\
2x(\delta-1) + (\delta-1)^2, & x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})
\end{cases}
\] 
But, when $x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta)$, 
\[
-1 \leq -1 + \delta \leq 2x + \delta < 1 - \delta \leq 1,
\]
and so
\[
-\delta \leq (-1 + \delta)\delta \leq (2x + \delta)\delta < (1-\delta)\delta \leq \delta.
\]
Also, when $x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})$ we  have $-\delta \leq 2x + \delta - 1 < \delta$, and consequently
\[
-\delta \leq -\delta(1-\delta) \leq (2x + \delta - 1)(1 - \delta) \leq \delta(1 - \delta) \leq \delta.
\]
%and consequenly
%\[
%-\delta \leq (2x + \delta - 1)(\delta - 1) \leq \delta.
%\]

@@

%The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.  The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.  
The remaining lemmas consider a fixed realisation $\{\Phi_n\}$, and we use the abbreviations $C_\epsilon = C_\epsilon(\{\Phi_n\})$ and $f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)$.  As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.  %This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% AOS,AOAS: If there are supplements please fill:
% \begin{supplement}[id=suppA]
%  \sname{Supplement A}
%  \stitle{Title}
%  \slink[url]{http://lib.stat.cmu.edu/aoas/???/???}
%  \sdescription{Some text}
% \end{supplement}

%%%%%%%%%%%%%%%%%%%%%%%%%%% END AOAP START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
% \documentclass[journal]{IEEEtran}

% \usepackage{mathbf-abbrevs}

% %\newcommand{\dealias}{\operatorname{dealias}}

% \input{defs}

% %\newcommand{\term}{\emph}
% %\renewcommand{\index}[1]{}

% %\title{Polynomial phase estimation by phase unwrapping 1: Identifiability and strong consistency}
% \title{Polynomial phase estimation by phase unwrapping}

% \author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
%     \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
% Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.  Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.   Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.  B. Moran is with the Department of Electrical Engineering and Computer
% Science, Melbourne Systems Lab, Dept of Elec \& Electronic Eng, Uni of Melbourne, Vic. 3010, Australia.}}
% % The paper headers 
% \markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}

% % make the title area

% \begin{document} 
% \maketitle

% \begin{abstract}
% Estimating the coefficients of a noisy polynomial phase signal is important in many fields including radar, biology and radio communications. One approach attempts to perform polynomial regression on the phase of the signal.  This is complicated by the fact that the phase is \emph{wrapped} modulo $2\pi$ and must be \emph{unwrapped} before regression can be performed. %A recent approach suggested by the authors is to perform the unwrapping in a least squares manner.  %It was shown how the resulting estimator could be posed as a \emph{nearest lattice point problem} in a specific lattice and could be solved using existing algorithms. 
% %It was shown by Monte Carlo simulation that this produces a remarkably accurate estimator.  
% In this paper we consider an estimator that performs phase unwrapping in a least squares manner.  We describe the asymptotic properties of this estimator, showing that it is strongly consistent and asymptotically normally distributed. %We hypothesise that the estimator produces very near maximum likelihood performance.
% \end{abstract}

% \begin{keywords}
% Polynomial phase signals, phase unwrapping, asymptotic properties, nearest lattice point problem
% \end{keywords}

% %\bibliographystyle{unsrt}

%%%%%%%%%%%%%%%%%%%%%%%%%%% END IEEE START %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

An obvious estimator of the unknown coefficients is the least squares estimator.  
%This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  
When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the least squares estimator is an effective approach, being both computationally efficient and statistically accurate~\cite{Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}. When $m \geq 2$ the computational complexity of the least squares estimator is large, and alternative estimators have been considered for this reason. These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995} and the high order phase function~\cite{Farquharson_another_poly_est_2005,Porat_asympt_HAF_DPT_1996}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.

@@

An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $o(N^{-k})$.  This is perhaps not surprising, since it is the same rate observed in polynomial regression.  However, asserting that convergence at this rate occurs in the polynomial phase setting is not trivial.  For this purpose we make use of an elementary result about the number of arithmetic progressions contained inside subsets of $\{1,2,\dots,N\}$~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.  %This proof technique appears to be novel, and may be useful in other problems.  
The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.  Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} and results from the literature on hyperplane arrangements~\cite{Chazelle_discrepency_method_2000,Matousek_lect_disc_geom_2002} become useful here.  We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particular other applications involving data that is `wrapped' in some sense.  Potential candidates are the phase wrapped images observed in modern radar and medical imaging devices such as synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}.

@@

A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^d \}
\]
where $\Bbf \in \reals^{n \times d}$ is an $n \times d$ matrix of rank $d$, called the generator matrix.  If $n = d$ the lattice is said to be full rank.  Lattices are discrete Abelian groups under vector addition.  They are subgroups of the Euclidean group $\reals^n$.  Lattices naturally give rise to tessellations of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.  One choice for a set of coset representatives is a fundamental parallelepiped; the parallelepiped generated by the columns of a generator matrix.  Another choice is based on the Voronoi cell those points from $\reals^n$ nearest (with respect to the Euclidean norm here) to the lattice point at the origin.  It is always possible to construct a rectangular set of representatives, as the next proposition will show.  We will use these rectangular regions for describing the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.  These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.

@@

In general there are infinitely many choices for the identifiable region. A natural choice is the Voronoi cell of $L_{m+1}$ used in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}. Another potential choice is a fundamental parallelepiped of $L_{m+1}$. In this paper we will use the rectangular set constructed using Proposition~\ref{prop:lattice-theory-constructing_a_rectangular_tesselating_region}. Observe that $\Pbf$ is upper triangular with $k$th diagonal element equal to $\tfrac{1}{k!}$.  So this rectangular set is
\begin{equation}\label{eq:rectangular_identifiable_region}
B = \prod_{k=0}^{m}\left[ -\frac{0.5}{k!}, \frac{0.5}{k!}  \right).
\end{equation}
We will make use of this set when deriving the statistical properties of the LSU estimator in the next section.

@@

Let $W = \{1,2,\dots, N\}$ and let $K$ be a subset of $W$.  For any integer $h$, let
\begin{equation} \label{eq:S(h,G)def} 
A(h,K) = \big\{ n \mid n + ih \in K \;\forall\; i \in \{0,1,\dots,m\} \big\}
\end{equation}
be the set containing all integers $n$ such that the arithmetic progression
\[
n, \,\, n + h, \,\, n + 2h, \,\, \dots, \,\, n + mh
\]
of length $m+1$ is contained in the subset $K$.  If $K$ is a small subset of $W$ then $A(h,K)$ might be empty. However, the next two lemmas and the following corollary will show that if $K$ is sufficiently large then it always contains at least one arithmetic progression (for all sufficiently small $h$) and therefore $A(h,K)$ is not empty. We do not wish to claim any novelty here, the study of arithmetic progressions within subsets of $W$ has a considerable history~\cite{Erdos_on_some_sequence_of_integers1936,Szemeredi_setint_no_k_arth1975,Gowers_new_proof2001}.  In particular, Gower's~\cite[Theorem 1.3]{Gowers_new_proof2001} gives a result far stronger than we require here.  Denote by $K \backslash \{r\}$ the set $K$ with the element $r$ removed.

@@

During the proof of strong consistency we made use of the fact that
\begin{equation}~\label{eq:SNVNunifmlln2}
\sup_{\lambdabf \in B}\sabs{S_N(\lambdabf) - V_N(\lambdabf)} \rightarrow 0
\end{equation}
almost surely as $N\rightarrow\infty$, where $V_N(\lambdabf) = \expect S_N(\lambdabf)$.  We prove this result here.  Put $D_N(\lambdabf) = S_N(\lambdabf) - V_N(\lambdabf)$.  Now $\sum_{N=1}^\infty \prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} < \infty$ for any $\epsilon > 0$ by Lemma~\ref{lem:vn}, and~\eqref{eq:SNVNunifmlln2} follows from the Borel-Cantelli lemma.   In what follows we use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.

\begin{lemma} \label{lem:vn} $\prob \left\{ \sup_{\lambdabf \in B}\abs{ D_N(\lambdabf) } > \epsilon \right\} = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 2$.
 \end{lemma}
\begin{proof}
Consider a rectangular grid of points spaced over the identifiable region $B$.  We use $\lambdabf[\rbf]$, where $\rbf \in \ints^{m+1}$, to denote the grid point 
 %\begin{align*}
\[ 
%\lambdabf[\rbf] = &\left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \right.\\ 
%&\dots, \; \frac{r_k}{N^{b+k}} - \frac{1}{2(k!)}, \; \dots \\
%&\left. \dots, \; \frac{r_m}{N^{b+m}} - \frac{1}{2(m!) }\right]
 \lambdabf[\rbf] = \left[  \frac{r_0}{N^{b}} - \frac{1}{2}, \; \frac{r_1}{N^{b+1}} - \frac{1}{2}, \; \dots, \; \frac{r_m}{m!N^{b+m}} - \frac{1}{2(m!) }\right]
\]
 %\end{align*}
 for some constant $b>0$.  Adjacent grid points are separated by $\tfrac{1}{N^b}$ in the zeroth coordinate, $\tfrac{1}{N^{b+1}}$ in the first coordinate and $\tfrac{1}{k!N^{b+k}}$ in the $k$th coordinate. Let
 \[
 B[\rbf]=\left\{  \xbf\in\reals^{m+1}  ; \frac{r_k}{N^{b+k}}\leq x_k + \frac{1}{2(k!)} < \frac{r_k + 1}{N^{b+k}} \right\}  .
 \]
and let $G$ be the finite set of grid points
\[
G = \left\{ \xbf \in \ints^{m+1} \mid x_k = 0,1,2\dots,N^{b+k}-1  \right\}.
\]
The total number of grid points is $|G| = N^{(m+1)(2b + m)/2}$, and the $B[\rbf]$ partition $B$, that is, $B = \cup_{\rbf \in G}B[\rbf]$.  Now 
\begin{align}
 \sup_{\lambdabf \in B}\sabs{ D_N(\lambdabf) }  &= \sup_{\rbf \in G}\sup_{ \lambdabf \in B[\rbf]}\sabs{ D_N(\lambdabf[\rbf]) + D_N(\lambdabf) - D_N(\lambdabf[\rbf])  } \nonumber \\
 &\leq \sup_{\rbf \in G}\sabs{ D_N(\lambdabf[\rbf])} + \sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\sabs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])  }. \label{eq:gridanddensespace}
 \end{align}
From Lemma~\ref{lem:supVjk} it will follow that $\prob\left\{   \sup_{\rbf \in G }\abs{ D_N(\lambdabf[\rbf])  } > \frac{\epsilon}{2} \right\} = O(e^{-c \epsilon^2 N})$
for any $\epsilon > 0$ and $c < 2$.  In Lemma~\ref{lem:supBVn} we show that
\[
\sup_{\rbf \in G}\sup_{\lambdabf\in B[\rbf]}\abs{ D_N(\lambdabf) - D_N(\lambdabf[\rbf])} < 2\frac{m+1}{N^b}.
\]
%that is, $D_N$ does not change much within any of the regions $B[\rbf]$ neighbouring the grid points.  
Combining these results with~\eqref{eq:gridanddensespace}, we obtain
\[
\prob\left(\sup_{\lambdabf \in B }\sabs{ D_N(\lambdabf)} > \frac{\epsilon}{2} +  \frac{2(m+1)}{N^b} \right) = O(e^{-c \epsilon^2 N}),
\]
and for sufficiently large $N$, we have $\epsilon/2 + \frac{2(m+1)}{N^b} < \epsilon$ completing the proof.  It remains to prove Lemmas~\ref{lem:supVjk} and~\ref{lem:supBVn}.
\end{proof}

@@

\begin{lemma}\label{lem:supVjk}
$\prob\cubr{   \sup_{\rbf \in G}\abs{ D_N(\lambdabf[\rbf])  } > \epsilon } = O(e^{-c \epsilon^2 N})$ for any $\epsilon > 0$ and $c < 8$.
\end{lemma}
 \begin{proof}
Fix $\lambdabf$ and write $D_N(\lambdabf) = \bar{Z} = \frac{1}{N}  \sum_{n=1}^{N}Z_{n}$, where
 \[
 Z_{n}=\fracpart{  \Phi_n+\sum_{k = 0}^{m}{\lambda_k n^k} }^{2} - \expect \fracpart{  \Phi_n + \sum_{k = 0}^{m}{\lambda_k n^k} }^{2}
 \]
 are independent with zero mean and $\abs{Z_n} \leq \tfrac{1}{4}$. It follows from Hoeffding's inequality~\cite{Hoeffding_inequality_1963} that,  $\prob\{ \abs{D_N(\lambdabf)} > \epsilon \} \leq 2e^{ - 8 \epsilon^2 N}$, and so,
 \begin{align*}
 \prob\cubr{  \sup_{\rbf\in G}\left\vert D_N(\lambdabf[\rbf]) \right\vert >\epsilon }  &\leq \sum_{\rbf \in G} \prob\scubr{  \left\vert D_N(\lambdabf[\rbf])  \right\vert >\epsilon } \nonumber \\
&= 2 |G| e^{ - 8 \epsilon^2 N} = O(e^{-c \epsilon^2 N}), \label{eq:betaprobconv}
\end{align*}
where $c$ is any real number less than $8$, since $|G| = N^{(m+1)(2b + m)/2}$ is polynomial in $N$.
\end{proof}

@@

\begin{lemma}\label{lem:boundedsquarefracparts}
$\fracpart{x}^2 - |\delta| \leq \fracpart{x + \delta}^2 \leq \fracpart{x}^2 + |\delta|$ for all $x, \delta \in \reals$.
\end{lemma}
\begin{proof}
Since $\abs{\delta} \leq \abs{n + \delta}$ for all $\delta \in [-\nicefrac{1}{2}, \nicefrac{1}{2})$ and $n \in \ints$, the result will follow if we can show that it holds when both $x$ and $\delta$ are in $[-\nicefrac{1}{2}, \nicefrac{1}{2})$.  Also, for reasons of symmetry, we need only show that it holds when $\delta \geq 0$.  Now
\[
\fracpart{x + \delta}^2 - x^2 = \begin{cases}
2x\delta + \delta^2, & x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta) \\
2x(\delta-1) + (\delta-1)^2, & x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})
\end{cases}
\] 
But, when $x \in [-\nicefrac{1}{2}, \nicefrac{1}{2} - \delta)$, 
\[
-1 \leq -1 + \delta \leq 2x + \delta < 1 - \delta \leq 1,
\]
and so
\[
-\delta \leq (-1 + \delta)\delta \leq (2x + \delta)\delta < (1-\delta)\delta \leq \delta.
\]
Also, when $x \in [\nicefrac{1}{2} - \delta, \nicefrac{1}{2})$ we  have $-\delta \leq 2x + \delta - 1 < \delta$, and consequently
\[
-\delta \leq -\delta(1-\delta) \leq (2x + \delta - 1)(1 - \delta) \leq \delta(1 - \delta) \leq \delta.
\]
%and consequenly
%\[
%-\delta \leq (2x + \delta - 1)(\delta - 1) \leq \delta.
%\]

@@

The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.  The proofs here follow those of Pollard~\cite{Pollard_asymp_empi_proc_1989}.  In the remaining lemmas we consider expectation conditional on $\{\Phi_n\}$ and treat $\{\Phi_n\}$ as a fixed realisation. We consequently use the abbreviations $C_\epsilon = C_\epsilon(\{\Phi_n\})$ and $f_{nN}(\psibf) = f_{nN}(\psibf, \Phi_n)$.  As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.  
%This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.

\begin{lemma}\label{lem:chaining2}
There exists a positive constant $K_1$ such that
\[
\expect_\Phi \sup_{\|\psibf\|_{\infty} < \epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
\]
\end{lemma}
\begin{proof}  Let $B_\epsilon = \{\xbf \in \reals \mid \|\xbf\|_\infty < \epsilon \}$.
%We are now required to prove that
%\[
%\expect_\Phi \sup_{\psibf \in B_\epsilon} \abs{ Z_N(\psibf) } \leq K_1 \sqrt{C_\epsilon}.
%\]
For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $\reals^{m+1}$ with the property that for every $\psibf \in B_\epsilon$ there exists some $\psibf^* \in T_\epsilon(k)$ such that the pseudometric
\[
d(\psibf, \psibf^*) = \sum_{n = 1}^N \big( f_{nN}(\psibf) - f_{nN}(\psibf^*) \big)^2 \leq 2^{-k} C_\epsilon N.
\]
We define $T_\epsilon(0)$ to contain a single point, the origin $\zerobf$.  Defined this way $T_\epsilon(0)$ satisfies the inequality above because $d(\psibf, \zerobf) = \sum_{n = 1}^N f_{nN}(\psibf)^2  \leq C_\epsilon N$ for all $\psibf \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
